### 2016-06-24 paper/algorithms.tex
changed subfigures and referencse to match siam style

\section{Inference Algorithm and Complexity Analysis}\label{sect:alg}
%
Based on the theoretical developments in the last section, we present below a goal-oriented inference algorithm that allows one to combine models of varying fidelity, while maintaining rigorous control of QoI error.
%
\subsection{Goal Oriented Inference Algorithm}
%
Just as error estimates can be used to guide mesh-refinement \cite{BecRann01}, the error estimate (\ref{eq:finErrExp}) can be localized to give elemental contributions and used to guide the division of the domain for a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. Algorithm~\ref{alg:refSeries} describes this approach. Note that the QoI error adjoint problem (\ref{eq:superAdjEq}) involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMF}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for

	supplementary adjoint $\Lambda_0$ (see Equation (\ref{eq:superAdjEq}))}
	\State{Compute QoI error estimate

@@

		$\quad\quad$supplementary adjoint $\Lambda_i$ (see Equation (\ref{eq:superAdjEq}))}
		\State{Compute QoI error estimate

@@

Algorithm~\ref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena, be a reduced-order model, or have a reduced parameter space. The two models could also correspond to two levels of mesh-refinement, though in this case the method described in~\cite{BecVex05} could be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead. The derived error estimate is not applicable to all models, however. The two models have to be expressed in a weak form, so this cannot be applied to, for example, a model of chemical reactions using kinetic Monte Carlo. We need some degree of compatibility between the two models; namely, we assume that $\Psi_{LF}$ will be in a space admissible to $\mathcal{M}'_{HF,\Psi}$, and that the QoI functional $I$ is applicable to both $(q_{HF},u_{HF})$ and $(q_{LF},u_{LF})$.
% 
\subsection{Complexity Analysis}\label{sect:alg_complexity}
%
We now derive the computational complexity of using Algorithm~\ref{alg:refSeries} to solve an inference problem, and compare it to extant strategies. Let the state, adjoint, and parameter variables each have $N$ degrees of freedom, and assume that an `all at once' strategy is used to solve the KKT system. The cost of single linear solve with such a system is $\orderof{3N}^{\gamma}$, where $\gamma$ depends on the linear solver used~\cite{}. The cost of solving for the auxiliary variables, a linear system, is $\orderof{3N}^{\gamma}$. \red{The parts of the supplementary adjoint corresponding to the primary and auxiliary variables can be solved separately, and thus the cost of computing the supplementary adjoint is also $\orderof{3N}^{\gamma}$.}

If one utilizes Algorithm~\ref{alg:refSeries} to solve the inference problem, with $T$ adaptive iterations the total computational cost is,
%
\begin{align}
\label{eq:cost_adapt}
C_{MF} &= \sum_{i=1}^{T}L_{i}(3N)^{\gamma} + 1(3N)^{\gamma} + 2(3N)^{\gamma} \nonumber \\
&= \sum_{i=1}^{T} (L_i + 3) (3N)^{\gamma}
\end{align}
% 
where $L_i$ is the number of nonlinear solver steps needed to solve the KKT system for the low or mixed fidelity model at each adaptive step. The corresponding cost for solving the high fidelity inference problem, with a nonadaptive algorithm, based on a continuation scheme is,
%
\begin{equation}
\label{eq:cost_nadapt}
C_{HF} = \sum_{i=1}^{C}K_{i}(3N)^{\gamma}
\end{equation}
% 
where $K_i$ is the number of nonlinear solver steps needed to solve the KKT system for the high fidelity model at each continuation step, and $C$ is the number of continuation steps.

Comparing Eq.~\eqref{eq:cost_adapt} and~\eqref{eq:cost_nadapt}, we have,
%
\begin{equation}
\label{eq:cost_compare_0}
C_{MF} < C_{HF} \implies \sum_{i=1}^{T} (L_i + 3) < \sum\limits_{i=1}^{C}K_{i}
\end{equation}
%
If the low and multi-fidelity models are linear, or nonlinear in a very small region, then $L_i \approx 1$, and we have the following condition,
%
\begin{equation}
\label{eq:cost_compare_1}
T < \frac{\sum\limits_{i=1}^{C}K_{i}}{4},
\end{equation}
% 
for the adaptive algorithm to be less costly than solving the high-fidelity inverse problem.
%The bound on $T$ will be strictest if Gaussian elimination is used, with $\gamma = 3$, giving $T < \frac{\sum\limits_{i=1}^{C}K_{i}}{10}$.
%

>>>>

\section{Inference Algorithm and Complexity Analysis}\label{sect:alg}
%
Based on the theoretical developments in the last section, we present below a goal-oriented inference algorithm that allows one to combine models of varying fidelity, while maintaining rigorous control of QoI error.
%
\subsection{Goal Oriented Inference Algorithm}
%
Just as error estimates can be used to guide mesh-refinement \cite{BecRann01}, the error estimate \cref{eq:finErrExp} can be localized to give elemental contributions and used to guide the division of the domain for a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. \Cref{alg:refSeries} describes this approach. Note that the QoI error adjoint problem \cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMF}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for

	supplementary adjoint $\Lambda_0$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

@@

		$\quad\quad$supplementary adjoint $\Lambda_i$ (see \cref{eq:superAdjEq})}
		\State{Compute QoI error estimate

@@

\Cref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena, be a reduced-order model, or have a reduced parameter space. The two models could also correspond to two levels of mesh-refinement, though in this case the method described in~\cite{BecVex05} could be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead. The derived error estimate is not applicable to all models, however. The two models have to be expressed in a weak form, so this cannot be applied to, for example, a model of chemical reactions using kinetic Monte Carlo. We need some degree of compatibility between the two models; namely, we assume that $\Psi_{LF}$ will be in a space admissible to $\mathcal{M}'_{HF,\Psi}$, and that the QoI functional $I$ is applicable to both $(q_{HF},u_{HF})$ and $(q_{LF},u_{LF})$.
% 
\subsection{Complexity Analysis}\label{sect:alg_complexity}
%
We now derive the computational complexity of using \cref{alg:refSeries} to solve an inference problem, and compare it to extant strategies. Let the state, adjoint, and parameter variables each have $N$ degrees of freedom, and assume that an `all at once' strategy is used to solve the KKT system. The cost of single linear solve with such a system is $\orderof{3N}^{\gamma}$, where $\gamma$ depends on the linear solver used~\cite{}. The cost of solving for the auxiliary variables, a linear system, is $\orderof{3N}^{\gamma}$. \red{The parts of the supplementary adjoint corresponding to the primary and auxiliary variables can be solved separately, and thus the cost of computing the supplementary adjoint is also $\orderof{3N}^{\gamma}$.}

If one utilizes \cref{alg:refSeries} to solve the inference problem, with $T$ adaptive iterations the total computational cost is,
%
\begin{align}
\label{eq:cost_adapt}
C_{MF} &= \sum_{i=1}^{T}L_{i}(3N)^{\gamma} + 1(3N)^{\gamma} + 2(3N)^{\gamma} \nonumber \\
&= \sum_{i=1}^{T} (L_i + 3) (3N)^{\gamma}
\end{align}
% 
where $L_i$ is the number of nonlinear solver steps needed to solve the KKT system for the low or mixed fidelity model at each adaptive step. The corresponding cost for solving the high fidelity inference problem, with a nonadaptive algorithm, based on a continuation scheme is,
%
\begin{equation}
\label{eq:cost_nadapt}
C_{HF} = \sum_{i=1}^{C}K_{i}(3N)^{\gamma}
\end{equation}
% 
where $K_i$ is the number of nonlinear solver steps needed to solve the KKT system for the high fidelity model at each continuation step, and $C$ is the number of continuation steps.

Comparing \cref{eq:cost_adapt} and \cref{eq:cost_nadapt}, we have,
%
\begin{equation}
\label{eq:cost_compare_0}
C_{MF} < C_{HF} \implies \sum_{i=1}^{T} (L_i + 3) < \sum\limits_{i=1}^{C}K_{i}
\end{equation}
%
If the low and multi-fidelity models are linear, or nonlinear in a very small region, then $L_i \approx 1$, and we have the following condition,
%
\begin{equation}
\label{eq:cost_compare_1}
T < \frac{\sum\limits_{i=1}^{C}K_{i}}{4},
\end{equation}
% 
for the adaptive algorithm to be less costly than solving the high-fidelity inverse problem.
%The bound on $T$ will be strictest if Gaussian elimination is used, with $\gamma = 3$, giving $T < \frac{\sum\limits_{i=1}^{C}K_{i}}{10}$.
%

### 2016-06-24 paper/intro.tex
changed subfigures and referencse to match siam style

The rest of the paper is organized as follows, Section~\ref{sect:form} presents the mathematical formulation and error analysis for the multi-model inverse problem. Then, in Section~\ref{sect:alg} we discuss the goal oriented inference algorithm and present a computational complexity analysis. Section~\ref{sect:numexp} shows the application of this algorithm to a model problem and a contaminant flow problem. Finally, we present conclusions and direction for future work in Section~\ref{sect:conc}.

>>>>

The rest of the paper is organized as follows, \cref{sect:form} presents the mathematical formulation and error analysis for the multi-model inverse problem. Then, in \cref{sect:alg} we discuss the goal oriented inference algorithm and present a computational complexity analysis. \cref{sect:numexp} shows the application of this algorithm to a model problem and a contaminant flow problem. Finally, we present conclusions and direction for future work in \cref{sect:conc}.

### 2016-06-24 paper/mathematical_formulation.tex
more paper edits
changed subfigures and referencse to match siam style

The unknown parameter $q$ can then be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
Thus, we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the constraints given by Equation (\ref{eq:weakForm}).

The constraints given by Eq.~\eqref{eq:weakForm} are typically models of physical processes or systems. However, a given physical system need not have a unique model that can describe it; there may be various models of different fidelities. This naturally introduces a hierarchy of models and, a tradeoff between fidelity and computational expense. Consider two models with which we infer parameters: a high-fidelity (HF) model and a low fidelity (LF) model. We then have a specific form of the inverse problem in Equation (\ref{eq:invOpt}) with this model,
\begin{equation}
\begin{array}{r@{}c}
\min\limits_{q_{HF},u_{HF}} & \quad J_{HF}(q_{HF},u_{HF})=\frac{1}{2}\|d-C_{HF}(u_{HF})\|_2^2 + R_{HF}(q_{HF}) \\ \textrm{s.t. }& \quad a_{HF}(u_{HF},q_{HF})(\phi_{HF})=\ell_{HF}(q_{HF})(\phi_{HF}),\quad\forall\phi_{HF}\in U_{HF},
\end{array}
\end{equation}
for which we have the Lagrangian,
\begin{multline}
\mathcal{L}_{HF}(q_{HF},u_{HF},z_{HF}) = \\
J_{HF}(q_{HF},u_{HF})-(a_{HF}(u_{HF},q_{HF})(z_{HF})-\ell_{HF}(q_{HF})(z_{HF})).
\end{multline}
The inverse problem formulation and Lagrangian with the lower fidelity model can be similarly defined, and be identified with subscript $LF$.

In the case of a goal-oriented inverse problem, the ultimate purpose of inferring the unknown parameters is to calculate some Quantity of Interest (QoI). Assuming a single scalar QoI, we denote this QoI by $I(q,u)$, where $I:Q\times U\to\R$ is a functional that maps the parameters and state to our QoI. Since our ultimate goal is to compute this QoI, the particular tradeoff we consider is the fidelity and resultant computational expense of the model we use, versus the error in the QoI. We thus seek to introduce the QoI functional $I$ into the inverse problem formulation, by introducing auxilary variables and additional adjoint equations. Then we use this formulation to derive an a posteriori error estimate for $I$, where the errors considered are those which occur due to the use of constraints other than those given by Eq.~\eqref{eq:invOpt_cons}.
%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse \\Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model. The main result is,
%
\begin{theorem}
\label{thm:error_estimate}
Following the notation established in section~\ref{sec:setup}, let the bilinear form $a:U \times U \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q\to\Reals$ be differentiable with respect to the parameter $q$, and the functional $I:Q\times U\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.

Consider the Lagrangian equation induced by Eq.~\eqref{eq:invOpt},
%
\begin{equation}
\label{eq:InvsOpt_lag}
\mathcal{L}(q,u,z)= J(q,u)-(a(u,q)(z)-\ell(q)(z)),
\end{equation}
%
where $z\in U$ is the adjoint. Denoting the primary variables as $\xi=(q,u,z)$, introduce corresponding auxiliary variables $\chi=(p,v,y)\in Q\times U\times U$. Let the augmented Lagrangian be defined as,
%
\begin{equation}
\label{eq:InvsOpt_auglag}
\mathcal{M}((q,u,z),(p,v,y)) = I(q,u) + \mathcal{L}_{quz}'(q,u,z)(p,v,y),
\end{equation}
%
where $\mathcal{L}_{quz}'(q,u,z)(p,v,y)$ denotes the Fr\'{e}chet derivative of the Lagrangian about the primary variables $(q,u,z)$, in the direction of the auxiliary variables $(p,v,y)$. Let $\Psi = (\xi_\Psi,\chi_\Psi)$ denote the stationary point of $\mathcal{M}$.

Denote by $\Psi_{HF}$ and $\Psi_{LF}$, stationary points of the high and low fidelity versions of Eq.~\eqref{eq:InvsOpt_auglag} and, consider the adjoint problem,
%
\begin{equation}
\label{eq:superAdjEq}
\mathcal{M}'_{HF,\Psi}(\Lambda;\Psi_{HF})(\Phi)=\mathcal{Q}(\Phi)=\mathcal M'_{HF}(\Psi_{LF})(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
\end{equation}
%
for the supplementary adjoint $\Lambda$. Then, the error in the Quantity of Interest $I$ is given by,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3).
\end{multline}
%
where $\mathcal{R}$ is a remainder term that is third-order in the error $e=\Psi_{HF}-\Psi_{LF}$.
\end{theorem}
%
\begin{proof}
%
Observe that,
%
\begin{equation}
\label{eq:MeqI}
\mathcal{M}(\Psi)=I(q,u),
\end{equation}
%
since taking variations of $\mathcal{M}$ with respect to the auxiliary variables gives that $\xi_\Psi$ is a stationary point of $\mathcal{L}$.

Extending the property in Equation (\ref{eq:MeqI}) to the augmented Lagrangians for the high and low fidelity models, we have,
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,
\begin{equation}
\label{eq:beckvex}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{,}
\end{equation}
Combining Equations (\ref{eq:repIwithM}) and (\ref{eq:beckvex}) we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{multline}

Further, the error in the output $\mathcal{Q}$ defined in Equation (\ref{eq:superAdjEq}) can be expressed as a dual-weighted residual,
\begin{equation}
\label{eq:adjOutErr}
\mathcal M'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})=-\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda).
\end{equation}

Combining Equations (\ref{eq:preadj}) and (\ref{eq:adjOutErr}), we have,
\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3). \nonumber
\end{multline}
%
which completes the proof.
\end{proof}
%
The above error estimate is general and the lower-fidelity model can also be a mixed-fidelity model that combines the high- and low-fidelity model. Given a low-fidelity model and a high-fidelity model, an intermediate, mixed-fidelity (MF) model can be formed by using the high-fidelity model in some parts of the domain, and the low-fidelity model in the rest of the domain.

>>>>

The unknown parameter $q$ can then be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
Thus, we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the constraints given by \cref{eq:weakForm}.

The constraints given by \cref{eq:weakForm} are typically models of physical processes or systems. However, a given physical system need not have a unique model that can describe it; there may be various models of different fidelities. This naturally introduces a hierarchy of models and, a tradeoff between fidelity and computational expense. Consider two models with which we infer parameters: a high-fidelity (HF) model and a low fidelity (LF) model. We then have a specific form of the inverse problem in \cref{eq:invOpt} with this model,
\begin{equation}
\begin{array}{r@{}c}
\min\limits_{q_{HF},u_{HF}} & \quad J_{HF}(q_{HF},u_{HF})=\frac{1}{2}\|d-C_{HF}(u_{HF})\|_2^2 + R_{HF}(q_{HF}) \\ \textrm{s.t. }& \quad a_{HF}(u_{HF},q_{HF})(\phi_{HF})=\ell_{HF}(q_{HF})(\phi_{HF}),\quad\forall\phi_{HF}\in U_{HF},
\end{array}
\end{equation}
for which we have the Lagrangian,
\begin{multline}
\mathcal{L}_{HF}(q_{HF},u_{HF},z_{HF}) = \\
J_{HF}(q_{HF},u_{HF})-(a_{HF}(u_{HF},q_{HF})(z_{HF})-\ell_{HF}(q_{HF})(z_{HF})).
\end{multline}
The inverse problem formulation and Lagrangian with the lower fidelity model can be similarly defined, and be identified with subscript $LF$.

In the case of a goal-oriented inverse problem, the ultimate purpose of inferring the unknown parameters is to calculate some Quantity of Interest (QoI). Assuming a single scalar QoI, we denote this QoI by $I(q,u)$, where $I:Q\times U\to\R$ is a functional that maps the parameters and state to our QoI. Since our ultimate goal is to compute this QoI, the particular tradeoff we consider is the fidelity and resultant computational expense of the model we use, versus the error in the QoI. We thus seek to introduce the QoI functional $I$ into the inverse problem formulation, by introducing auxilary variables and additional adjoint equations. Then we use this formulation to derive an a posteriori error estimate for $I$, where the errors considered are those which occur due to the use of constraints other than those given by \cref{eq:invOpt_cons}.
%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model. The main result is,
%
\begin{theorem}
\label{thm:error_estimate}
Following the notation established in \cref{sec:setup}, let the bilinear form $a:U \times U \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q\to\Reals$ be differentiable with respect to the parameter $q$, and the functional $I:Q\times U\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.

Consider the Lagrangian equation induced by \cref{eq:invOpt},
%
\begin{equation}
\label{eq:InvsOpt_lag}
\mathcal{L}(q,u,z)= J(q,u)-(a(u,q)(z)-\ell(q)(z)),
\end{equation}
%
where $z\in U$ is the adjoint. Denoting the primary variables as $\xi=(q,u,z)$, introduce corresponding auxiliary variables $\chi=(p,v,y)\in Q\times U\times U$. Let the augmented Lagrangian be defined as,
%
\begin{equation}
\label{eq:InvsOpt_auglag}
\mathcal{M}((q,u,z),(p,v,y)) = I(q,u) + \mathcal{L}_{quz}'(q,u,z)(p,v,y),
\end{equation}
%
where $\mathcal{L}_{quz}'(q,u,z)(p,v,y)$ denotes the Fr\'{e}chet derivative of the Lagrangian about the primary variables $(q,u,z)$, in the direction of the auxiliary variables $(p,v,y)$. Let $\Psi = (\xi_\Psi,\chi_\Psi)$ denote the stationary point of $\mathcal{M}$.

Denote by $\Psi_{HF}$ and $\Psi_{LF}$, stationary points of the high and low fidelity versions of \cref{eq:InvsOpt_auglag} and, consider the adjoint problem,
%
\begin{equation}
\label{eq:superAdjEq}
\mathcal{M}'_{HF,\Psi}(\Lambda;\Psi_{HF})(\Phi)=\mathcal{Q}(\Phi)=\mathcal M'_{HF}(\Psi_{LF})(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
\end{equation}
%
for the supplementary adjoint $\Lambda$. Then, the error in the Quantity of Interest $I$ is given by,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3).
\end{multline}
%
where $\mathcal{R}$ is a remainder term that is third-order in the error $e=\Psi_{HF}-\Psi_{LF}$.
\end{theorem}
%
\begin{proof}
%
Observe that,
%
\begin{equation}
\label{eq:MeqI}
\mathcal{M}(\Psi)=I(q,u),
\end{equation}
%
since taking variations of $\mathcal{M}$ with respect to the auxiliary variables gives that $\xi_\Psi$ is a stationary point of $\mathcal{L}$.

Extending the property in \cref{eq:MeqI} to the augmented Lagrangians for the high and low fidelity models, we have,
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,
\begin{equation}
\label{eq:beckvex}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{,}
\end{equation}
Combining \cref{eq:repIwithM} and \cref{eq:beckvex} we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{multline}

Further, the error in the output $\mathcal{Q}$ defined in \cref{eq:superAdjEq} can be expressed as a dual-weighted residual,
\begin{equation}
\label{eq:adjOutErr}
\mathcal M'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})=-\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda).
\end{equation}

Combining \cref{eq:preadj} and \cref{eq:adjOutErr}, we have,
\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3). \nonumber
\end{multline}
%
which completes the proof.
\end{proof}
%
The above error estimate is general and the lower-fidelity model can also be a mixed-fidelity model that combines the high- and low-fidelity model. Given a low-fidelity model and a high-fidelity model, an intermediate, mixed-fidelity (MF) model can be formed by using the high-fidelity model in some parts of the domain, and the low-fidelity model in the rest of the domain.
